\singlespacing \vspace{-1.6cm} \LARGE Covariation among reproductive traits in flowering plants determine interactions with floral visitors

### 2022-06-28 Manuscript/Lanuza_et_al_2022.tex
edits

Plants have an enormous variation of life history strategies and trait combinations. However, there is strong evidence that there are evolutionary and physiological constraints that limit the number of plant ecological strategies. Despite recent advances in the understanding of plant trait variation, little is known about reproductive trait correlations and how these affect plant life-history strategies and shape interactions with floral visitors. Here, we investigate the reproductive spectrum of flowering plants and how this drive interactions with floral visitors using a dataset of 16 reproductive traits for 1,506 plant species. We found that over 50\% and 70\% of all trait variation was explained by the first two and three reproductive axes, respectively. Specifically, the first axis indicated the presence of a negative correlation between flower number and flower size; the second axis showed a positive correlation between pollinator dependence and floral display; and the third axis indicated a negative correlation between style length and pollinator dependence. These different reproductive axes determined important differences in the interaction level among floral visitor guilds but were insufficient to fully capture plant-pollinator associations. Our study shows the major reproductive trait correlations and highlights their relevance to understand plant-pollinator interactions in a global context.

@@

Flowering plants have an astonishing diversity of floral structures {[}\protect\hyperlink{ref-barrett2002}{1},\protect\hyperlink{ref-schiestl2013}{2}{]} that shape plant-pollinator associations {[}\protect\hyperlink{ref-fenster2004}{3},\protect\hyperlink{ref-dellinger2020}{4}{]}. However, not all reproductive trait combinations are possible due to evolutionary and ecological constraints {[}\protect\hyperlink{ref-stearns1989}{5},\protect\hyperlink{ref-agrawal2020}{6}{]}. Despite the recent advances in the theoretical and empirical understanding of the macroecological correlations between plant reproductive traits {[}\protect\hyperlink{ref-salguero2016}{7}--\protect\hyperlink{ref-roddy2021}{10}{]}, their study is often limited to a handful number of reproductive traits that are rarely studied jointly. Thus, in order to progress towards a comprehensive understanding of the plant reproductive spectrum of trait variation as done recently for other morphological and physiological plant traits {[}\protect\hyperlink{ref-wright2004}{11}--\protect\hyperlink{ref-laughlin2021}{15}{]}, there is a need to acquire a multitrait perspective with broad geographical coverage. Importantly, the characterization of the reproductive spectrum can help to further understand the different plant ecological strategies {[}\protect\hyperlink{ref-agrawal2020}{6}{]} and improve our knowledge on plant-pollinator associations {[}\protect\hyperlink{ref-roddy2021}{10}{]}.

Although there is an increasing number of macroecological studies that investigate plant reproductive traits {[}\protect\hyperlink{ref-baude2016}{16}--\protect\hyperlink{ref-moeller2017}{19}{]}, we still have poor understanding of how reproductive traits drive interactions with floral visitors at large ecological scales {[}\protect\hyperlink{ref-salguero2016}{7},\protect\hyperlink{ref-sargent2008}{20}--\protect\hyperlink{ref-ruger2018}{22}{]}. In addition, the pollination system of a great number of plant species remains unexplored and is still unclear how specific key reproductive traits like mating or compatibility system influence plant-pollinator associations {[}\protect\hyperlink{ref-tur2013}{23},\protect\hyperlink{ref-devaux2014}{24}{]}. Interestingly, the use of trait-based approaches {[}\protect\hyperlink{ref-fenster2004}{3},\protect\hyperlink{ref-rosas2014}{25}{]} and trait-matching analyses {[}\protect\hyperlink{ref-stang2009}{26},\protect\hyperlink{ref-bartomeus2016}{27}{]} has shown to be of great importance when exploring the drivers of plant-pollinator interactions. For example, plant traits can define species' network roles {[}e.g., specialists vs generalists \protect\hyperlink{ref-tur2013}{23}{]} and plant species that occupy reproductive trait space extremes are more likely to exhibit higher levels of specialisation and be more reliant on the trait-matching with pollinators {[}\protect\hyperlink{ref-junker2013}{28},\protect\hyperlink{ref-coux2016}{29}{]}. Indeed, morphological matching between plants and floral visitors often determines plant-pollinator interactions, and can thus strongly influence interaction network structure {[}\protect\hyperlink{ref-stang2009}{26},\protect\hyperlink{ref-ibanez2012}{30}{]}. Because the species' morphology can determine the species' functional role in the pollination network and the combination of traits has shown to increase the predictive power of the network of interactions {[}\protect\hyperlink{ref-eklof2013}{31}{]}, an interesting novel approach is to investigate how traits in the multidimensional trait space determine species interaction patterns {[}see \protect\hyperlink{ref-dehling2016}{32}{]}. Thus, by exploring the plant reproductive spectrum of trait variation is possible to delimit the different reproductive strategies and explore how these are associated with the different floral visitors.

With the recent availability of large trait databases {[}e.g., TRY \protect\hyperlink{ref-kattge2011}{33},~COMPADRE \protect\hyperlink{ref-salguero2015}{34}{]}, plant ecological strategies are being increasingly examined, and are facilitating the identification of global patterns and constraints in plant form and function {[}\protect\hyperlink{ref-salguero2016}{7},\protect\hyperlink{ref-diaz2016}{13},\protect\hyperlink{ref-bruelheide2018}{35},\protect\hyperlink{ref-carmona2021}{36}{]}. However, most studies with a multitrait perspective have focused on trait correlations from the leaf {[}\protect\hyperlink{ref-wright2004}{11}{]}, wood {[}\protect\hyperlink{ref-chave2009}{12}{]}, or root {[}\protect\hyperlink{ref-laughlin2021}{15}{]} related traits with little or no attention given to reproductive traits {[}\protect\hyperlink{ref-roddy2021}{10},\protect\hyperlink{ref-evojtko2020}{37}{]}. Despite the lack of an holistic view that depict reproductive trait covariation patterns, there are widely recognized reproductive trait associations between pair of traits such as the negative correlation between flower size and flower number {[}\protect\hyperlink{ref-sargent2007}{38},\protect\hyperlink{ref-kettle2011}{39}{]}, the positive association between flower size and outcrossing rate {[}\protect\hyperlink{ref-goodwillie2010}{40}{]} or the positive association between outcrossing rate and lifespan {[}\protect\hyperlink{ref-moeller2017}{19},\protect\hyperlink{ref-barrett2003}{41}{]}. Although these different trait correlations (and others) have recently allowed to progress towards a conceptual framework that integrates the different floral trait relationships {[}\protect\hyperlink{ref-roddy2021}{10}{]}, we still lack empirical evidence that investigates jointly how these different reproductive trait associations shape plant life history strategies.

@@

We selected 28 studies from 18 different countries that constituted a total of 64 plant-pollinator networks (see table S1 and figure S1). These studies recorded plant-pollinator interactions in natural systems and were selected so that we had broad geographical representation across different biological communities. Although these studies differ in sampling effort and methodology, all studies provided information about plant-pollinator interactions (weighted and non-weighted), which we used to build a database of plant species that are likely to be animal pollinated. Many of these networks are freely available either as published studies {[}e.g., \protect\hyperlink{ref-carvalheiro2014}{42},\protect\hyperlink{ref-olesen2007}{43},\protect\hyperlink{ref-fortuna2010}{44}{]} or available in online archives {[}e.g., Mangal \protect\hyperlink{ref-poisot2016}{45},~The Web of Life \protect\hyperlink{ref-fortuna2010}{44}{]}. In total, our network dataset constituted 60 weighted (number of visits) and 4 unweighted (presence-absence of the interaction) networks, each sampled at a unique location and year, as well as eight meta-webs where interactions were pooled across several locations and multiple years.

@@

All species names, genera, families and orders were retrieved and standardized from the taxonomy data sources NCBI (\url{https://www.ncbi.nlm.nih.gov/taxonomy}) for plants and ITIS (\url{https://www.itis.gov/}) for pollinators, using the R package \emph{taxize} {[}\protect\hyperlink{ref-chamberlain2020}{46}{]}. We filled the `not found' searches manually using \url{http://www.theplantlist.org/} and \url{http://www.mobot.org/} for plants and \url{http://www.catalogueoflife.org/} for floral visitors.

@@

\caption{\label{tab:unnamed-chunk-1}\textbf{Table 1.} Quantitative and categorical traits used in this study.}
\centering
\fontsize{12}{14}\selectfont
\begin{tabular}[t]{ll}
\toprule
\textbf{Traits} & \textbf{Numerical range/Categories}\\
\midrule
\addlinespace[0.3em]
\multicolumn{2}{l}{\textbf{Quantitative}}\\
\hspace{1em}Plant height & 0.01-30m\\
\hspace{1em}Flower width & 0.50-205mm\\
\hspace{1em}Flower length & 0.20-195mm\\
\hspace{1em}Inflorescence width & 0.80-300mm\\
\hspace{1em}Style length & 0-138mm\\
\hspace{1em}Ovules per flower & 1-25 x $10^3$\\
\hspace{1em}Flowers per plant & 1-18 x $10^4$\\
\hspace{1em}Microliters of nectar & 0-160$\mu$l\\
\hspace{1em}Milligrams of nectar & 0-6.1mg\\
\hspace{1em}Nectar concentration & 0-100$\%$\\
\hspace{1em}Pollen grains per flower & 13-2 x $10^8$\\
\hspace{1em}Autonomous selfing (fruit set) & 0-100$\%$\\
\addlinespace[0.3em]
\multicolumn{2}{l}{\textbf{Categorical}}\\
\hspace{1em}Lifepan & Short-lived and perennial\\
\hspace{1em}Life form & Herb, shrub and tree\\
\hspace{1em}Flower shape & \makecell[l]{Brush, campanulate, capitulum,\\ open, papilionaceous and tube}\\
\hspace{1em}Flower symmetry & Actinomorphic and zygomorphic\\
\hspace{1em}Nectar & Presence and absence\\
\hspace{1em}Compatibility system & \makecell[l]{Self-incompatible, partially self-compabtile \\ and self-compatible}\\
\hspace{1em}Breeding system & Hermaphrodite, monoecious and dioecious\\
\bottomrule
\end{tabular}
\end{table}

@@

We calculated the phylogenetic distance between different plant species using the function \emph{get\_tree} from the package \emph{rtrees} (\url{https://github.com/daijiang/rtrees}), which downloads phylogenetic distances from the extended R implementation of the Open Tree of Life {[}\protect\hyperlink{ref-smith2018}{47},\protect\hyperlink{ref-jin2019}{48}{]}.

@@

To avoid the loss of relevant ecological information, we imputed trait missing values with the help of the function \emph{missForest} {[}\protect\hyperlink{ref-stekhoven2012}{49}{]} which allows imputation of datasets with continuous and categorical variables. We accounted for the phylogenetic distance among species on the imputation process by including the eigenvectors of a principal component analysis of the phylogenetic distance (PCoA) which has been shown to improve the performance of \emph{missForest} {[}\protect\hyperlink{ref-penone2014}{50}{]}. To extract the eigenvectors, we used the function \emph{PVRdecomp} from the package \emph{PVR} {[}\protect\hyperlink{ref-santos2018}{51}{]} based on a previous conceptual framework that considers phylogenetic eigenvectors {[}\protect\hyperlink{ref-diniz-filho2012}{52}{]}. We conducted two different imputations, one for the full set of species (1,506 species, 5.79\% of missing values) excluding nectar and pollen traits because of the high percentage of missing values (figure S2) and a second one for the subset of species with data for pollen per flower and microliters of nectar (755 species, 8.01\% of missing values). To corroborate that our imputation of missing values did not affect our results, we evaluated the reproductive spectrum (see section below) with and without missing values and we found consistent similar results on both analyses (Fig S3 and Fig S4).

@@

We explored the association between the different quantitative plant traits with a phylogenetically informed Principal Component Analysis (pPCA). We did not include the quantitative variables of flower length and inflorescence width because they were highly and moderately correlated to flower width respectively (Pearson's correlation = 0.72, \emph{P} \textless{} 0.01 and Pearson's correlation = 0.36, \emph{P} \textless{} 0.01), and thus we avoided overemphasizing flower size on the spectrum of trait variation. Prior to the analyses, we excluded outliers and standardized the data. Due to the high sensitivity of dimensionality reduction to outliers {[}\protect\hyperlink{ref-serneels2008}{53},\protect\hyperlink{ref-legendre2012}{54}{]}, we excluded values outside the 2.5th--97.5th percentile range, and thus our final dataset had 1,236 species. Then, we log transformed the variables to reduce the influence of outliers and z-transformed (X= 0, SD=1) so that all variables were within the same numerical range as indicated for principal component analysis {[}\protect\hyperlink{ref-legendre2012}{54}{]}. Although qualitative traits were not included in the dimensionality reduction analysis, we also investigated the statistical association of the different qualitative traits with the main axes of trait variation with the help of an Anova and a Tukey test. We performed the pPCA using the function \emph{phyl.pca} from the package \emph{phytools} {[}\protect\hyperlink{ref-revell2012}{55}{]} with the method lambda (\(\lambda\)) that calculates the phylogenetic correlation between 0 (phylogenetic independence) and 1 (shared evolutionary history) and we implemented the mode covariance because values for each variables were on the same scale following transformation {[}\protect\hyperlink{ref-abdi2010}{56}{]}. Finally, we conducted an additional phylogenetic informed principal component analysis for the subset of species with pollen and nectar quantity. For this, we included all quantitative traits considered in the main pPCA plus pollen grains and microlitres of nectar per flower.

@@

We calculated the phylogenetic signal of the different quantitative traits on the imputed dataset with the full set of species (N = 1,506) with the package \emph{phytools} {[}\protect\hyperlink{ref-revell2012}{55}{]} and we used Pagel's \(\lambda\) as a measurement of the phylogenetic signal. However, for pollen and nectar traits, phylogenetic signal was calculated only on the subset of species that had quantitative information for these traits (N = 755).

@@

First, we investigated how the different groups of floral visitors interacted along the main axes of reproductive trait variation (see below `visitation patterns' section) with the help of qualitative and quantitative information of plant-pollinator interactions. For this, we used as qualitative information the binary version of the networks (presence-absence of interaction) that assumes equal weight across interactions and as quantitative information the number of visits of floral visitors to individual flowers that accounts for the intensity of the interaction. Although floral visitors are not always pollinators and number of visits does not consider each pollinator species efficiency {[}\protect\hyperlink{ref-ballantyne2015}{57}{]}, the number of visits can provide valuable information of the contribution of floral visitors to pollination {[}\protect\hyperlink{ref-vazquez2005}{58},\protect\hyperlink{ref-vazquez2012}{59}{]}. Second, we investigated how the main axes of trait variation and individual traits influence plant species' functional roles in the pollination network using a set of complementary interaction network metrics: number of visits, normalized degree and specialization (see below `plant species network roles' section).

Analyses were conducted on the subset of 60 weighted networks sampled in a unique flowering season and site, which included 556 plant and 1,126 pollinator species. In total, our network dataset (excluding meta-webs and non-weighted networks) included 2,256 interactions of bees with plants, 1,768 non-syrphid-Diptera interactions, 845 syrphids interactions, 437 Lepidoptera interactions, 432 Coleoptera interactions and 362 non-bee-Hymenoptera interactions. Sampling methods varied across networks but this was accounted for in analyses by considering them in the random effects of the modelling process. All analyses were conducted in R version \emph{4.0.3}.

@@

We used Bayesian modelling (see below for details) to explore the effect of floral visitor groups and the main axes of trait variation (pPCA with imputed dataset) on both qualitative and quantitative floral interactions per plant species. For this, we divided floral visitors into six main guilds that differ in life form, behaviour and are likely to play a similar ecological role: (i) bees (Hymenoptera-Anthophila), (ii) non-bee-Hymenoptera (Hymenoptera-non-Anthophila), (iii) syrphids (Diptera-Syrphidae), (iv) non-syrphid-Diptera (Diptera-non-Syrphidae), (v) Lepidoptera and (vi) Coleoptera. Moreover, because the guild of bees was the most represented group with 2,256 records and had the highest frequency of visits, we also explored presence-absence of the interaction and number of visits of the main bee families (Andrenidae, Apidae, Colletidae, Halictidae and Megachilidae) on the trait space. In addition, we found that \emph{Apis mellifera} was the floral visitor with the largest proportion of records counted (7.55\% of the total). This finding is consistent with previous research showing that \emph{A. mellifera} was the most frequent floral visitor in a similar dataset of 80 plant-pollinator networks in natural ecosystems {[}\protect\hyperlink{ref-hung2018}{60}{]}. Hence, to control for the effect of \emph{A. mellifera} on the observed visitation patterns of bees, we conducted an analogous analysis with presence-absence of the interaction and number of visits excluding \emph{A. mellifera}. We found that \emph{A. mellifera}, was partly driving some of the observed trends on PC1 (figure S5). However, we did not detect major differences on PC2 and PC3.

We implemented Bayesian generalized linear mixed models using the R package \emph{brms} {[}\protect\hyperlink{ref-burkner2017}{61}{]}. We modelled presence-absence of observed interactions and number of visits as a function of the main axes of plant trait variation and their interactions with floral visitor guilds (e.g., number of visits \textasciitilde{} PC1 x FGs + PC2 x FGs + PC3 x FGs). Because we were interested in possible differences in the visitation patterns among floral visitors groups to plants with different strategies, we included interactions between the main axes of trait variation (PC1, PC2 and PC3) and the different floral visitor guilds. We added a nested random effect of networks nested within the study system to capture the variation in networks among studies and within networks. Moreover, we included the phylogenetic covariance matrix as a random factor due to the possible shared evolutionary histories of species and therefore lack of independence across them. We specified for presence-absence of interaction and number of visits a Bernoulli and a zero inflated negative binomial distribution, respectively. The models were run with 3,000 iterations with previous 1,000 warm up iterations and with non or very weakly informative informative priors from the brm function so they have neglible influence on the results {[}\protect\hyperlink{ref-burkner2017}{61}{]}. We set delta (\(\Delta\)) to 0.99 to avoid divergent transitions and visualized the posterior predictive checks with the function \emph{pp\_check} using the \emph{bayesplot} package {[}\protect\hyperlink{ref-gabry2019}{62}{]}.

@@

{\centering \includegraphics[width=0.95\linewidth]{output/figures/unnamed-chunk-2-1}

@@

\leavevmode\vadjust pre{\hypertarget{ref-wright2004}{}}%
\CSLLeftMargin{11. }
\CSLRightInline{Wright IJ \emph{et al.} 2004 The worldwide leaf economics spectrum. \emph{Nature} \textbf{428}, 821--827. (doi:\href{https://doi.org/10.1038/nature02403}{10.1038/nature02403})}

>>>>

Plants have an enormous variation of life history strategies and trait combinations. However, there is strong evidence that there are evolutionary and physiological constraints that limit the number of plant ecological strategies. Despite recent advances in the understanding of plant trait correlations, reproductive traits are often not considered despite its key role in shaping both plant life-history strategies and plants interactions with floral visitors. Here, we investigate the reproductive spectrum of flowering plants and how this drive the establishment of interactions with their floral visitors using a dataset of 16 reproductive traits for 1,506 plant species. We found that over 50\% of all trait variation was explained by the first two reproductive axes. Specifically, the first axis indicated the presence of a negative correlation between flower number and flower size while the second axis showed a positive correlation between pollinator dependence and floral display. These different reproductive axes determined important differences in the interaction level among floral visitor guilds but were insufficient to fully capture plant-pollinator associations. Our study shows the main reproductive trait correlations of flowering plants and highlights their relevance to understand plant-pollinator interactions in a global context.

@@

Flowering plants have an astonishing diversity of floral structures {[}\protect\hyperlink{ref-barrett2002}{1},\protect\hyperlink{ref-schiestl2013}{2}{]} that shape their interactions with animal pollinators {[}\protect\hyperlink{ref-fenster2004}{3},\protect\hyperlink{ref-dellinger2020}{4}{]}. The variety of reproductive strategies followed by different plant species, from specialized reliance on a single pollinator, to full self-pollination, is incredibly rich. However, not all reproductive trait combinations are possible due to evolutionary and ecological constraints {[}\protect\hyperlink{ref-stearns1989}{5},\protect\hyperlink{ref-agrawal2020}{6}{]}. Despite the recent advances in the theoretical and empirical understanding of the macroecological correlations between plant reproductive traits {[}\protect\hyperlink{ref-salguero2016}{7}--\protect\hyperlink{ref-roddy2021}{10}{]}, their study is often limited to a few number of reproductive traits that are rarely studied in combination. Thus, in order to progress towards a comprehensive understanding of the plant reproductive spectrum of trait variation as done recently for other root {[}\protect\hyperlink{ref-laughlin2021}{11}{]}, leaf {[}\protect\hyperlink{ref-wright2004}{12}{]} or wood {[}\protect\hyperlink{ref-chave2009}{13}{]} related traits, there is a need to acquire a multitrait perspective with broad geographical coverage. Importantly, the characterization of the plant reproductive spectrum can help to further understand the different plant ecological strategies {[}\protect\hyperlink{ref-agrawal2020}{6}{]} and improve our knowledge on plant-pollinator associations {[}\protect\hyperlink{ref-roddy2021}{10}{]}.

With the recent availability of large trait databases {[}e.g., TRY \protect\hyperlink{ref-kattge2011}{14},~COMPADRE \protect\hyperlink{ref-salguero2015}{15}{]}, plant ecological strategies are being increasingly examined, and are facilitating the identification of global patterns and constraints in plant form and function {[}\protect\hyperlink{ref-salguero2016}{7},\protect\hyperlink{ref-diaz2016}{16}--\protect\hyperlink{ref-carmona2021}{18}{]}. However, most studies with a multitrait perspective have overlooked plant reproductive traits {[}\protect\hyperlink{ref-roddy2021}{10},\protect\hyperlink{ref-evojtko2020}{19}{]} and focused on trait correlations from other plant organs. Despite the lack of an holistic view that depict reproductive trait covariation patterns, there are widely recognized reproductive trait associations between pair of traits such as the negative correlation between flower size and flower number {[}\protect\hyperlink{ref-sargent2007}{20},\protect\hyperlink{ref-kettle2011}{21}{]}, the positive association between flower size and outcrossing rate {[}\protect\hyperlink{ref-goodwillie2010}{22}{]} or the positive association between outcrossing rate and lifespan {[}\protect\hyperlink{ref-barrett2003}{23},\protect\hyperlink{ref-moeller2017}{24}{]}. Although these different trait correlations (and others) have recently allowed to progress towards a conceptual framework that integrates the different floral trait relationships {[}\protect\hyperlink{ref-roddy2021}{10}{]}, we still lack empirical evidence that investigates jointly how these different reproductive trait associations shape plant life history strategies.

Despite this increasing number of macroecological studies that investigate specific plant reproductive traits {[}\protect\hyperlink{ref-moeller2017}{24}--\protect\hyperlink{ref-grossenbacher2017}{27}{]}, we still have poor understanding of how reproductive traits drive interactions with floral visitors at large ecological scales {[}\protect\hyperlink{ref-salguero2016}{7},\protect\hyperlink{ref-sargent2008}{28}--\protect\hyperlink{ref-ruger2018}{30}{]}. The use of trait-based approaches {[}\protect\hyperlink{ref-fenster2004}{3},\protect\hyperlink{ref-rosas2014}{31}{]} and trait-matching analyses {[}\protect\hyperlink{ref-stang2009}{32},\protect\hyperlink{ref-bartomeus2016}{33}{]} has shown to be of great importance when exploring the drivers of plant-pollinator interactions. For example, plant traits can define species' network roles {[}e.g., specialists vs generalists \protect\hyperlink{ref-tur2013}{34}{]} and plant species that occupy reproductive trait space extremes are more likely to exhibit higher levels of specialisation and be more reliant on the trait-matching with pollinators {[}\protect\hyperlink{ref-junker2013}{35},\protect\hyperlink{ref-coux2016}{36}{]}. Indeed, morphological matching between plants and floral visitors often determines plant-pollinator interactions, and can thus strongly influence interaction network structure {[}\protect\hyperlink{ref-stang2009}{32},\protect\hyperlink{ref-ibanez2012}{37}{]}. However, is still unclear how specific key reproductive traits like mating or compatibility system influence plant-pollinator associations {[}\protect\hyperlink{ref-tur2013}{34},\protect\hyperlink{ref-devaux2014}{38}{]}. Indeed, the pollination system of a great number of plant species remains unexplored {[}\protect\hyperlink{ref-dellinger2020}{4}{]}. Because both the species' morphology and mating system can determine the species' functional role in the pollination network and the combination of several traits has shown to increase the predictive power of the structure of ecological networks {[}\protect\hyperlink{ref-eklof2013}{39}{]}, an interesting novel approach is to investigate how a complete set of reproductive traits in the multidimensional trait space determine species interaction patterns {[}see \protect\hyperlink{ref-dehling2016}{40} for a morphological approach{]}. Thus, by exploring the plant reproductive spectrum of trait variation is possible to delimit the different reproductive strategies and explore how these are associated with the different floral visitors.

@@

We selected 28 studies from 18 different countries that constituted a total of 64 plant-pollinator networks (see table S1 and figure S1). These studies recorded plant-pollinator interactions in natural systems and were selected so that we had broad geographical representation across different biological communities. Although these studies differ in sampling effort and methodology, all studies provided information about plant-pollinator interactions (weighted and non-weighted), which we used to build a database of plant species that are likely to be animal pollinated. Many of these networks are freely available either as published studies {[}e.g., \protect\hyperlink{ref-carvalheiro2014}{41},\protect\hyperlink{ref-olesen2007}{42},\protect\hyperlink{ref-fortuna2010}{43}{]} or available in online archives {[}e.g., Mangal \protect\hyperlink{ref-poisot2016}{44},~The Web of Life \protect\hyperlink{ref-fortuna2010}{43}{]}. In total, our network dataset constituted 60 weighted (number of visits) and 4 unweighted (presence-absence of the interaction) networks, each sampled at a unique location and year, as well as eight meta-webs where interactions were pooled across several locations and multiple years.

@@

All species names, genera, families and orders were retrieved and standardized from the taxonomy data sources NCBI (\url{https://www.ncbi.nlm.nih.gov/taxonomy}) for plants and ITIS (\url{https://www.itis.gov/}) for pollinators, using the R package \emph{taxize} {[}\protect\hyperlink{ref-chamberlain2020}{45}{]}. We filled the `not found' searches manually using \url{http://www.theplantlist.org/} and \url{http://www.mobot.org/} for plants and \url{http://www.catalogueoflife.org/} for floral visitors.

@@

\caption{\label{tab:unnamed-chunk-1}\textbf{Table 1.} Quantitative and categorical traits used in this study.}
\centering
\fontsize{12}{14}\selectfont
\begin{tabular}[t]{ll}
\toprule
\textbf{Traits} & \textbf{Numerical range/Categories}\\
\midrule
\addlinespace[0.3em]
\multicolumn{2}{l}{\textbf{Quantitative}}\\
\hspace{1em}Plant height & 0.01-30m\\
\hspace{1em}Flower width & 0.50-205mm\\
\hspace{1em}Flower length & 0.20-195mm\\
\hspace{1em}Inflorescence width & 0.80-300mm\\
\hspace{1em}Style length & 0-138mm\\
\hspace{1em}Ovules per flower & 1-25000\\
\hspace{1em}Flowers per plant & 1-1.8 x $10^5$\\
\hspace{1em}Microliters of nectar & 0-160$\mu$l\\
\hspace{1em}Milligrams of nectar & 0-6.1mg\\
\hspace{1em}Nectar concentration & 0-100$\%$\\
\hspace{1em}Pollen grains per flower & 13-2 x $10^8$\\
\hspace{1em}Autonomous selfing (fruit set) & 0-100$\%$\\
\addlinespace[0.3em]
\multicolumn{2}{l}{\textbf{Categorical}}\\
\hspace{1em}Lifepan & Short-lived and perennial\\
\hspace{1em}Life form & Herb, shrub and tree\\
\hspace{1em}Flower shape & \makecell[l]{Brush, campanulate, capitulum,\\ open, papilionaceous and tube}\\
\hspace{1em}Flower symmetry & Actinomorphic and zygomorphic\\
\hspace{1em}Nectar & Presence and absence\\
\hspace{1em}Compatibility system & \makecell[l]{Self-incompatible, partially self-compabtile \\ and self-compatible}\\
\hspace{1em}Breeding system & Hermaphrodite, monoecious and dioecious\\
\bottomrule
\end{tabular}
\end{table}

@@

We calculated the phylogenetic distance between different plant species using the function \emph{get\_tree} from the package \emph{rtrees} (\url{https://github.com/daijiang/rtrees}), which downloads phylogenetic distances from the extended R implementation of the Open Tree of Life {[}\protect\hyperlink{ref-smith2018}{46},\protect\hyperlink{ref-jin2019}{47}{]}.

@@

To avoid the loss of relevant ecological information, we imputed trait missing values with the help of the function \emph{missForest} {[}\protect\hyperlink{ref-stekhoven2012}{48}{]} which allows imputation of datasets with continuous and categorical variables. We accounted for the phylogenetic distance among species on the imputation process by including the eigenvectors of a principal component analysis of the phylogenetic distance (PCoA) which has been shown to improve the performance of \emph{missForest} {[}\protect\hyperlink{ref-penone2014}{49}{]}. To extract the eigenvectors, we used the function \emph{PVRdecomp} from the package \emph{PVR} {[}\protect\hyperlink{ref-santos2018}{50}{]} based on a previous conceptual framework that considers phylogenetic eigenvectors {[}\protect\hyperlink{ref-diniz-filho2012}{51}{]}. We conducted two different imputations, one for the full set of species (1,506 species, 5.79\% of missing values) excluding nectar and pollen traits because of the high percentage of missing values (figure S2) and a second one for the subset of species with data for pollen per flower and microliters of nectar (755 species, 8.01\% of missing values). To corroborate that our imputation of missing values did not affect our results, we evaluated the reproductive spectrum (see section below) with and without imputed values and we found consistent similar results on both analyses (Fig S3 and Fig S4).

@@

We explored the association between the different quantitative plant traits with a phylogenetically informed Principal Component Analysis (pPCA). We did not include the quantitative variables of flower length and inflorescence width because they were highly and moderately correlated to flower width respectively (Pearson's correlation = 0.72, \emph{P} \textless{} 0.01 and Pearson's correlation = 0.36, \emph{P} \textless{} 0.01), and thus we avoided overemphasizing flower size on the spectrum of trait variation. Prior to the analyses, we excluded outliers and standardized the data. Due to the high sensitivity of dimensionality reduction to outliers {[}\protect\hyperlink{ref-serneels2008}{52},\protect\hyperlink{ref-legendre2012}{53}{]}, we excluded values outside the 2.5th--97.5th percentile range, and thus our final dataset had 1,236 species. Then, we log transformed the variables to reduce the influence of outliers and z-transformed (X= 0, SD=1) so that all variables were within the same numerical range as recomended for principal component analysis {[}\protect\hyperlink{ref-legendre2012}{53}{]}. Although qualitative traits were not included in the dimensionality reduction analysis, we also investigated the statistical association of the different qualitative traits with the main axes of trait variation with the help of an Anova and a Tukey test. We performed the pPCA using the function \emph{phyl.pca} from the package \emph{phytools} {[}\protect\hyperlink{ref-revell2012}{54}{]} with the method lambda (\(\lambda\)) that calculates the phylogenetic correlation between 0 (phylogenetic independence) and 1 (shared evolutionary history) and we implemented the mode covariance because values for each variables were on the same scale following transformation {[}\protect\hyperlink{ref-abdi2010}{55}{]}. Finally, we conducted an additional phylogenetic informed principal component analysis for the subset of species with pollen and nectar quantity. For this, we included all quantitative traits considered in the main pPCA plus pollen grains and microlitres of nectar per flower.

@@

We calculated the phylogenetic signal of the different quantitative traits on the imputed dataset with the full set of species (N = 1,506) with the package \emph{phytools} {[}\protect\hyperlink{ref-revell2012}{54}{]} and we used Pagel's \(\lambda\) as a measurement of the phylogenetic signal. However, for pollen and nectar traits, phylogenetic signal was calculated only on the subset of species that had quantitative information for these traits (N = 755).

@@

@@

First, we used Bayesian modelling (see below for details) to investigate how the different groups of floral visitors interacted along the main axes of reproductive trait variation with the help of qualitative and quantitative information of plant-pollinator interactions. For this, we used as qualitative information the binary version of the networks (presence-absence of interaction) that assumes equal weight across interactions and as quantitative information the number of visits of floral visitors to individual flowers that accounts for the intensity of the interaction. Although floral visitors are not always pollinators and number of visits does not consider each pollinator species efficiency {[}\protect\hyperlink{ref-ballantyne2015}{56}{]}, the number of visits can provide valuable information of the contribution of floral visitors to pollination {[}\protect\hyperlink{ref-vazquez2005}{57},\protect\hyperlink{ref-vazquez2012}{58}{]}. For this, we divided floral visitors into six main guilds that differ in life form, behaviour and are likely to play a similar ecological role: (i) bees (Hymenoptera-Anthophila), (ii) non-bee-Hymenoptera (Hymenoptera-non-Anthophila), (iii) syrphids (Diptera-Syrphidae), (iv) non-syrphid-Diptera (Diptera-non-Syrphidae), (v) Lepidoptera and (vi) Coleoptera. Moreover, because the guild of bees was the most represented group with 2,256 records and had the highest frequency of visits, we also explored presence-absence of the interaction and number of visits of the main bee families (Andrenidae, Apidae, Colletidae, Halictidae and Megachilidae) on the trait space. In addition, we found that \emph{Apis mellifera} was the floral visitor with the largest proportion of records counted (7.55\% of the total). This finding is consistent with previous research showing that \emph{A. mellifera} was the most frequent floral visitor in a similar dataset of 80 plant-pollinator networks in natural ecosystems {[}\protect\hyperlink{ref-hung2018}{59}{]}. Hence, to control for the effect of \emph{A. mellifera} on the observed visitation patterns of bees, we conducted an analogous analysis with presence-absence of the interaction and number of visits excluding \emph{A. mellifera}. We found that \emph{A. mellifera}, was partly driving some of the observed trends on PC1 (figure S5). However, we did not detect major differences on PC2 and PC3.

We implemented Bayesian generalized linear mixed models using the R package \emph{brms} {[}\protect\hyperlink{ref-burkner2017}{60}{]}. We modelled presence-absence of observed interactions and number of visits as a function of the main axes of plant trait variation and their interactions with floral visitor guilds (e.g., number of visits \textasciitilde{} PC1 x FGs + PC2 x FGs + PC3 x FGs). Because we were interested in possible differences in the visitation patterns among floral visitors groups to plants with different strategies, we included interactions between the main axes of trait variation (PC1, PC2 and PC3) and the different floral visitor guilds. We added a nested random effect of networks nested within the study system to capture the variation in networks among studies and within networks. Moreover, we included the phylogenetic covariance matrix as a random factor due to the possible shared evolutionary histories of species and therefore lack of independence across them. We specified for presence-absence of interaction and number of visits a Bernoulli and a zero inflated negative binomial distribution, respectively. The models were run with 3,000 iterations with previous 1,000 warm up iterations and with non or very weakly informative informative priors from the brm function so they have neglible influence on the results {[}\protect\hyperlink{ref-burkner2017}{60}{]}. We set delta (\(\Delta\)) to 0.99 to avoid divergent transitions and visualized the posterior predictive checks with the function \emph{pp\_check} using the \emph{bayesplot} package {[}\protect\hyperlink{ref-gabry2019}{61}{]}.

@@

{\centering \includegraphics[width=0.85\linewidth]{output/figures/unnamed-chunk-2-1}

@@

\leavevmode\vadjust pre{\hypertarget{ref-laughlin2021}{}}%
\CSLLeftMargin{11. }
\CSLRightInline{Laughlin DC \emph{et al.} 2021 Root traits explain plant species distributions along climatic gradients yet challenge the nature of ecological trade-offs. \emph{Nat. Ecol. Evol.}, 1--12. (doi:\href{https://doi.org/10.1038/s41559-021-01471-7}{10.1038/s41559-021-01471-7})}

\leavevmode\vadjust pre{\hypertarget{ref-wright2004}{}}%
\CSLLeftMargin{12. }
\CSLRightInline{Wright IJ \emph{et al.} 2004 The worldwide leaf economics spectrum. \emph{Nature} \textbf{428}, 821--827. (doi:\href{https://doi.org/10.1038/nature02403}{10.1038/nature02403})}
